---
title: "Collaborative Filtering with R"
output: html_notebook
---

***
**Objective**: *Build a collaborative filtering model to recommend movies using the [MovieLense dataset](https://grouplens.org/datasets/movielens/)*

***

Load Libraries
```{r}
library(recommenderlab)
library(data.table)
```
## 1. Data Preparation

Load data and convert to ratings matrix. Using dcast to create matrix from data frame and then cast it to matrix, ratingmatrix
```{r}
ratings <- read.csv("ratings.csv")
ratingmat <- dcast(ratings, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingmat <- as.matrix(ratingmat[,-1])
ratingmat <- as(ratingmat, "realRatingMatrix")
```

###Visualize data
```{r}
image(ratingmat[1:100,1:100])
```
Look at number of raings per user
```{r}
hist(rowCounts(ratingmat))
```

Number of ratings per movie

```{r}
hist(colCounts(ratingmat))
```


Subset Movies that have been rated more than 50 times and Users who have rated atleast 25 movies

```{r}
Ratings50M <- ratingmat[rowCounts(ratingmat)>50,colCounts(ratingmat)>25]
```

```{r}
Ratings50M
```

## 2. Modeling
Create a recommender system using Userbased CF
```{r}
rmodel <- Recommender(Ratings50M, method = "UBCF")
```
Get recommendations for User 1
```{r}
recommendations <- predict(rmodel, Ratings50M[1:2], n=5)
as(recommendations, "list")[[1]]
```

Load Movies
```{r}
movies <- read.csv("movies.csv")
```
List movies from recommendations
```{r}
movies[movies$movieId %in% as(recommendations, "list")[[1]],1:2]
```

